Relative parabolic factorizations of Chevalley groups

### 2015-04-04 congrsubgr-boundgen.tex
Some references and small fixes

@@

@@

\textsc{Case 2:} $\alpha<\beta$. Since $\mathbb{Z}[\sfrac{1}{p}]/I$ is finite, there exists $k>0$ such that $p^k\equiv 1\pmod I$. One can choose $k>\beta-\alpha$. Then $k+\alpha>-k+\beta$ and
\[ g\cdot h\left(p^k\right) =
\begin{pmatrix} p^\alpha a & p^\beta bm \\ * & * \end{pmatrix}
\begin{pmatrix} p^k & 0 \\ 0 & p^{-k} \end{pmatrix}=
\begin{pmatrix} p^{k+\alpha} a & p^{-k+\beta} bm \\ * & * \end{pmatrix}.
\]
One can apply Case 1 for the latter, so
$g=x_{21}z_{12}x_{12}x_{21}h^{-1}$,
and $x_{12}h^{-1}$ can be processed in the same way, as in Lemma \ref{lemma:srRI1}:
$g=x_{21}z_{12}x_{12}\cdot z_{12}x_{21}x_{12}$.
\end{proof}
Let $\mathcal{O}_S$ be a Dedekind ring of arithmetic type, defined by a finite set $S$ of places of some global field, and $I$ an ideal in $\mathcal{O}_S$. It is shown in [?] that $E(\Phi,I)$ has finite width with respect to $\{x_\alpha(s)\mid\alpha\in\Phi,\ s\in I\}$. From this one immediately gets the bounded generation for $\G(\Phi,R,I)$.

@@

$G(I,I)$ denotes the Vaserstein's congruence subgroup
\[ G(I,I)=\left\{ \begin{pmatrix}1+a & b \\ c & 1+d\end{pmatrix}\in\SL(2,\mathcal{O}_S,I)\;\middle|\; a,d\in I^2,\ b,c\in I \right\}. \]
Note that $g\cdot z_{21}(a,1)\in G(I,I)$. Indeed,
\[
\begin{pmatrix}
1+a & b \\ c & 1+d
\end{pmatrix} \xmapsto{\cdot z_{21}(a,1)}
\begin{pmatrix}
1+ba-a^2 & b-a-ba-a^2 \\ c+a+ad-ac & 1+bc-ac
\end{pmatrix} \in G(I,I).
\]
Now for a matrix $g'=\begin{psmallmatrix}1+a & b \\ c & 1+d\end{psmallmatrix}\in G(I,I)$ the matrix $x_{21}(-c)\cdot g'\cdot x_{12}(-b)\in\G(2,\mathcal{O}_S,I^2)$. Since for root systems $\Phi\neq\rC_\ell$ with $\rk\Phi\geqslant2$ one has $\E(\Phi,\mathcal{O}_S,I^2)\leqslant\E(\Phi,I)$, we are done for all groups but symplectic.

>>>>

The following lemma is a corollary of Theorems~5.7 and 5.8 of \cite{LenMorStePrimitiveRoot}.

@@

One can modify Lemma~6 of \cite{VavSmSuUnitrEng} (see also \cite{VseUnitrZ1p}) in the following way:

@@

\textsc{Case 2:} $\alpha<\beta$. Since $\mathbb{Z}[\sfrac{1}{p}]/I$ is finite, there exists $k>0$ such that $p^k\equiv 1\pmod I$. One can choose $k>\beta-\alpha$. Then $k+\alpha>-k+\beta$ and
\[ g\cdot h\left(p^k\right) =
\begin{pmatrix} p^\alpha a & p^\beta bm \\ * & * \end{pmatrix}
\begin{pmatrix} p^k & 0 \\ 0 & p^{-k} \end{pmatrix}=
\begin{pmatrix} p^{k+\alpha} a & p^{-k+\beta} bm \\ * & * \end{pmatrix}.
\]
One can apply Case 1 for the latter, so
$g=x_{21}z_{12}x_{12}x_{21}h^{-1}$,
and $x_{12}h^{-1}$ can be processed in the same way, as in Lemma \ref{lemma:srRI1}:
$g=x_{21}z_{12}x_{12}\cdot z_{12}x_{21}x_{12}$.
\end{proof}
Let $\mathcal{O}_S$ be a Dedekind ring of arithmetic type, defined by a finite set $S$ of places of some global field, and $I$ an ideal in $\mathcal{O}_S$. It is shown in \cite{TavgenThesis} that $E(\Phi,I)$ has finite width with respect to $\{x_\alpha(s)\mid\alpha\in\Phi,\ s\in I\}$. From this one immediately gets the bounded generation for $\G(\Phi,R,I)$.

@@

Define the Vaserstein's congruence subgroup as
\[ G(I,I)=\left\{ \begin{pmatrix}1+a & b \\ c & 1+d\end{pmatrix}\in\SL(2,\mathcal{O}_S)\;\middle|\; a,d\in I^2,\ b,c\in I \right\}. \]
Note that $g\cdot z_{21}(a,1)\in G(I,I)$. Indeed,
\[
\begin{pmatrix}
1+a & b \\ c & 1+d
\end{pmatrix} \xmapsto{\cdot z_{21}(a,1)}
\begin{pmatrix}
1+ba-a^2 & b-a-ba-a^2 \\ c+a+ad-ac & 1+bc-ac
\end{pmatrix} \in G(I,I).
\]
Now for a matrix $g'=\begin{psmallmatrix}1+a & b \\ c & 1+d\end{psmallmatrix}\in G(I,I)$ the matrix $x_{21}(-c)\cdot g'\cdot x_{12}(-b)\in\G(2,\mathcal{O}_S,I^2)$. Since for root systems $\Phi\neq\rC_\ell$ with $\rk\Phi\geqslant2$ one has $\E(\Phi,\mathcal{O}_S,I^2)\leqslant\E(\Phi,I)$, we are done for all groups but symplectic.

### 2015-04-04 paper.tex
Given definition of asr
Make the bounded generation section readable
Some references and small fixes

\documentclass[12pt]{amsart}
\usepackage{amscd, amsmath, verbatim, enumitem, graphicx, amssymb, mathtools, xfrac, cite}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage[backend=biber, citestyle=numeric-comp, natbib=true, sortlocale=en_US, url=false, doi=false, eprint=true, maxbibnames=4]{biblatex}            
\usepackage[all]{xy}

@@

\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\G}{G}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\EP}{EP}
\DeclareMathOperator{\U}{U}
\DeclareMathOperator{\X}{X}
\DeclareMathOperator{\Z}{Z}
\DeclareMathOperator{\M}{M}
\DeclareMathOperator{\sr}{sr}
\DeclareMathOperator{\Rad}{Rad}
\DeclareMathOperator{\SR}{SR}
\DeclareMathOperator{\asr}{asr}
%TODO: Replace Um with Ums/Umd where possible
\DeclareMathOperator{\Um}{Um}
\DeclareMathOperator{\Ums}{Ums}
\DeclareMathOperator{\Umd}{Umd}
\DeclareMathOperator{\rk}{rk}
\newcommand{\rA}{\mathsf{A}}
\newcommand{\rB}{\mathsf{B}}
\newcommand{\rC}{\mathsf{C}} 
\newcommand{\rD}{\mathsf{D}} 
\newcommand{\rE}{\mathsf{E}}

@@

\begin{abstract} \end{abstract}

@@

\section {Introduction}\label{sec:intro}

@@

\section {Preliminaries}\label{sec:prelim}
For any collection of subsets $H_1,\ldots, H_n$ of a group $G$ we denote by $H_1\ldots H_n$ their Minkowski set-product,
i.\,e. the set consisting of arbitrary products $h_1\ldots h_n$ of elements $h_i\in H_i$. In particular, the equality
$G = H_1\cdot\ldots\cdot H_n$ means that every element $g\in G$ can be presented as a product $h_1\ldots h_n$ for $h_i\in H_i$.

>>>>

\documentclass[12pt]{amsart}
\usepackage{amscd, amsmath, verbatim, enumitem, graphicx, amssymb, mathtools, xfrac}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[backend=biber, citestyle=numeric-comp, natbib=true, sortlocale=en_US, url=false, doi=false, eprint=true, maxbibnames=4]{biblatex}            
\usepackage[all]{xy}

@@

\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\G}{G}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\SL}{SL}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\EP}{EP}
\DeclareMathOperator{\U}{U}
\DeclareMathOperator{\X}{X}
\DeclareMathOperator{\Z}{Z}
\DeclareMathOperator{\M}{M}
\DeclareMathOperator{\SR}{SR}
\DeclareMathOperator{\sr}{sr}
\DeclareMathOperator{\Rad}{Rad}
\DeclareMathOperator{\Max}{Max}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\ASR}{ASR}
\DeclareMathOperator{\asr}{asr}
%TODO: Deprecate Um -- Use Ums/Umd whenever possible
\DeclareMathOperator{\Um}{Um}
\DeclareMathOperator{\Ums}{Ums}
\DeclareMathOperator{\Umd}{Umd}
\DeclareMathOperator{\rk}{rk}
\newcommand{\rA}{\mathsf{A}}
\newcommand{\rB}{\mathsf{B}}
\newcommand{\rC}{\mathsf{C}} 
\newcommand{\rD}{\mathsf{D}} 
\newcommand{\rE}{\mathsf{E}}

@@

%\begin{abstract} \end{abstract}

@@

\section{Introduction}\label{sec:intro}

@@

\section{Preliminaries}\label{sec:prelim}
For any collection of subsets $H_1,\ldots, H_n$ of a group $G$ we denote by $H_1\ldots H_n$ their Minkowski set-product,
i.\,e. the set consisting of arbitrary products $h_1\ldots h_n$ of elements $h_i\in H_i$. In particular, the equality
$G = H_1\cdot\ldots\cdot H_n$ means that every element $g\in G$ can be presented as a product $h_1\ldots h_n$ for $h_i\in H_i$.
